---
title: "R For Data Science - Homework 1"
author: "Francisco Grisanti - PSID 1565858 - Date 2/21/2019"
output: 
    tufte::tufte_html: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = FALSE}

## this are the libraries used for my homework
## uploaded by the library() function
library(ISLR)
library(tufte)
library(tufterhandout)
library(ggplot2)
library(ggthemes)
```

```{r echo=FALSE}

### we assign the variable obs to the dataframe Weekly

obs <- Weekly
```

#Weekly Dataset

## Weekly percentage returns for the S&P 500 stock index between 1990 and 2010. 

## A data frame with 1089 observations on the following 9 variables.

#

### Year    - The year that the observation was recorded

### Lag1    - Percentage return for previous week

### Lag2    - Percentage return for 2 weeks previous

### Lag3    - Percentage return for 3 weeks previous

### Lag4    - Percentage return for 4 weeks previous

### Lag5    - Percentage return for 5 weeks previous

### Volume  - Volume of shares traded (average number of daily shares traded in billions)

### Today   - Percentage return for this week

### Direction - A factor with levels Down and Up indicating whether the market had a positive or negative 

## The dataset is organized this way (see subset) and multiple observations are gathered across weeks of years 1990-2010

```{r echo=FALSE}
knitr::kable(
  #we use a subset to display the user the form of the dataframe
  obs[1:5,1:9],caption = 'A subset of Weekly dataset'
  )
```

## The summary of all variables and observations 


```{r echo=FALSE}

## we call the function sumary() to find all the max,min,etc about out different variables of our dataframe

knitr::kable(
        summary(obs[1:9]), caption = 'Summary statistics of all features(min, max, mean, median, 1st Q, 3rd Q)'
                     )
```

## We can see clearly on the numerically & graphically that the market went up 605 days and Down 485 days

```{r echo=FALSE}
knitr::kable(summary(obs$Direction), caption = 'Total count of days the market was up or down')

# we use ggplot to plot a bar chart of the data 

p <- ggplot(obs, aes(x = obs$Direction)) + geom_bar() 
p + theme_tufte() + xlab("Direction of Market") + ylab("Total count")

```

##Studiying the volume across time we notice an exponential increment 

^['We see that up until 2005 the increment in the volume of trades is lineal, but after that breaking point the rise is exponential. Wee also a rare decrease around 2008, year in which US suffered a big recesion']

```{r echo=FALSE}

## we use ggplot to create a two variable plot with geom_jitter aesthetics
p <- ggplot(obs, aes(x = obs$Year, y=obs$Volume )) + geom_jitter(color="gray22") 
p + theme_tufte() + xlab("Year") + ylab("Volume") 


```

##Studiying the return of the stocks (Today's variable) across time

^['We see that the 2000 & 2008 crisis are represented by a distribution comparatively more tall than the others, and with more outliers']

```{r echo=FALSE}

## we use ggplot to create a boxplot across all year factors

obs$Year <- factor(obs$Year,labels= c("1990",'1991','1992','1993','1994','1995','1996','1997','1998','1999','2000','2001','2002','2003','2004','2005','2006','2007','2008','2009','2010'))

p <- ggplot(obs, aes(x = obs$Year, y=obs$Today )) + geom_boxplot() 

p + theme_tufte() + xlab("Year") + ylab("")

```





